\section{Virtual Machine}

\textbf{Underlying our proposed language is a
minimal virtual machine that follows the RISC philosophy}~\cite{patterson1985reduced}.
This means that individual machine instructions
perform basic units of work rather than elaborate functions.
It also implies a register-based ``load-store" architecture,
which limits memory access to \texttt{LOAD} and \texttt{STORE} instructions.
Finally, it implies that the memory addressing modes of the \texttt{LOAD} and
\texttt{STORE} instructions are few and simple. This architecture was chosen
for our virtual machine because the operation of
RISC machines is conceptually simple, making them
appealing targets for compilation, emulation, and visualization.

Like physical machines, the virtual machine processes
linear sequences of instructions and uses jump or call instructions
to alter the flow of control. This is in contrast with ``structured programming"
where code is organized in blocks or as a tree structure.
As in the case with x86 and ARM CPUs, \textbf{conditional branching depends on flags in
a status register}. For portability reasons, the status register
of the virtual machine cannot be read or modified directly.

Also like many physical machines, \textbf{our virtual machine requires that data be
properly aligned}. No automatic alignment is provided.
In addition, \textbf{the endianness of the virtual machine is not defined.}
Layout of multi-byte primitive data types depends on the underlying machine.

Because the virtual machine is very similar to modern hardware,
\textbf{machine code generated from our language should not require a
runtime library}. An exception to this is necessary when a fundamental operation
(e.g., integer division) is not available as an instruction (or as a short
sequence of instructions) on a particular CPU architecture (e.g., ARM).
In this case, a function call to a runtime library is needed.
This library could be from a third-party (e.g., libgcc).

\subsection{Registers}

Physical machines have a limited number of registers, and this number
can vary significantly from architecture to architecture.
Rather than commit to a certain number of registers, \textbf{our virtual
machine provides an unlimited number of registers}. These registers
may or may not map directly to actual machine registers.

Although single static assignment form is convenient for performing
optimizing code transformations, we chose instead to follow physical machine
register semantics. Thus, \textbf{our virtual machine allows
registers to be assigned a value multiple times.}

Modern CPUs segregate register banks by data type.
For example, modern machines have registers for memory addresses
that are separate from the registers for floating point values.
But there are also several cases where data types will share the same bank.
For example, registers for memory addresses are also used for signed integer
values. This is also the case for floating point values and integer SIMD vectors.

To avoid this complication,
\textbf{our virtual machine provides a separate register bank for
each primitive data type}.
It is left to the machine code generator to select the
appropriate register bank of the actual machine in each case.

\subsection{Machine Word Width}
\label{sec:word}

Physical machines have a natural ``word" data type that is wide enough
for working with memory address values. For this same purpose,
our virtual machine has word registers. But, because the width of this data type
varies from physical machine to physical machine, \textbf{the width of word registers in
our virtual machine is undefined.}

Unfortunately, it is now unclear how much
memory should be allocated for storage of word values.
We resolve this problem by specifiying that
\textbf{word values stored in memory are assumed to occupy 64-bits.}
This leaves enough room for word values of all modern machines.
Optimizing storage of word values in memory on 32-bit machines
is not a priority.

\subsection{Instructions}

Some instruction sets do not always separate source operands from
destination operands (i.e., a source operand must double as the destination
in some cases).
This can make the machine a more difficult compilation target.
Thus, \textbf{source and destination operands are separate} in our
virtual machine instructions.
Of course, it is possible to specify a register as both a source
and destination operand in the same instruction.

Rather than define or adopt a specific ABI, virtual machine \textbf{pseudo-instructions
provide platform-specific functionality}.
For example, allocating call stack memory is provided by a pseudo-instruction
because the task requires knowing the direction that the stack grows.
Also, function calling is provided by a pseudo-instruction because
it requires knowing how arguments are passed on the call stack.
An important advantage of implementing the platform ABI is that it allows for interfacing
with machine code that was not compiled via our virtual machine abstraction.

A description of the instruction set of the virtual machine
is given in Section 4.
